### 2015-12-29 feedback.tex
add copy of feedback.tex from exante and compile

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
Feedback on Jan. 10 draft
\end{center}

Chad Bown, Jan. 22, 2015
\begin{itemize}
	\item What it's missing: intersection with reality
		\begin{itemize}
			\item Anecdotes
			\item How policy is used
				\begin{itemize}
					\item I thought that escape clause implications \textit{were} this
				\end{itemize}
			\item Something like how GH predicts export taxes but it doesn't work that way in the real world
			\item Well articulated question / paradox
		\end{itemize}
	\item Look at ``Theory of Managed Trade'' model motivation (BS)
	\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
		\begin{itemize}
			\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
			\item You can have one good idea in a paper, the rest get lost
		\end{itemize}
	\item How did Devashish structure his AER paper: the argument, the model?
		\begin{itemize}
			\item This paper is structured differently than the norm---is there a good reason to do it?
			\item Don't want to challenge what people are used to unless I really need to
		\end{itemize}
	\item In general, how do AER papers \textit{look}?
	\item Intro doesn't reflect body of paper
	\item What kinds of empirical questions might we ask with this model?
		\begin{itemize}
			\item What would be different because we use this model?
		\end{itemize}
\end{itemize}

\end{document}