### 2016-08-18 game-tree.tex
more mods to game-tree.tex, include so it can compile standalone for now

%\begin{comment}
\documentclass[10pt]{article}

\addtolength{\textwidth}{1.3in}
\addtolength{\oddsidemargin}{-.65in} %left margin
\addtolength{\evensidemargin}{-.65in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\usepackage{setspace}
\doublespacing

\usepackage{titling}

\usepackage[pdftex]{graphicx}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont,pdfsync}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{tikz}

% disables chapter, section and subsection numbering
\setcounter{secnumdepth}{-1}

\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
\usepackage[hang,flushmargin]{footmisc}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}

\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\bte}{\bm{\tau^E}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}

\begin{document}
%\end{comment}

\tikzset{
% Two node styles for game trees: solid and hollow
solid node/.style={circle,draw,inner sep=1.5,fill=black},
hollow node/.style={circle,draw,inner sep=1.5,fill=white}
%non node/.style={circle,draw,inner sep=.1,fill=black}
}
\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
% Specify spacing for each level of the tree
\tikzstyle{level 1}=[level distance=10mm,sibling distance=8mm]
\tikzstyle{level 2}=[level distance=8mm,sibling distance=8mm]
% The Tree
\node(0)[solid node,label=above:{$G,G^*$}]{} 
child{node(1)[solid node,label=below:{$0$}]{}}
child{[white] node(2)[solid node,xshift=12,label=above:{L}]{}  
	child{[black] node(4)[solid node,label=below:{$0$}]{}}
	child{[white] node(5)[hollow node,xshift=12]{} 
	 %child[sibling distance=11mm]{[black] node(22)[solid node]{}  }
	 child{[black] node(22)[solid node]{}  }
	 child{[white] node(24)[solid node,xshift=12]{}
		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
		child{[white] node(8)[solid node,xshift=12]{}  
			child{[black] node(10)[solid node]{}  }
			child{[white] node(11)[solid node,xshift=14]{} }
			child{[black] node(12)[solid node]{}  }
			%edge from parent node[black, xshift=20]{$\ta$} 
			}
		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
		}
	 child{[black] node(23)[solid node]{}  }
	 edge from parent node[black, xshift=20]{$e_b$} 	
	 }
	child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
	edge from parent node[black, xshift=23]{$\bta$} 
	}
child{node(3)[solid node,label=above:{$\infty$}]{}
}
;
% information set
    \draw[solid,bend right](1)to(3);
		\draw[solid,bend right](4)to(6);
		\draw[solid,bend right](7)to(9);
		\draw[solid,bend right](10)to(12);
		\draw[solid,bend right](22)to(23);
% labels
		\node[above,xshift=3,yshift=5]at(2){$L$}; 
		%\node[above,xshift=1,yshift=5]at(23){$N$}; 
		\node[above,xshift=3,yshift=5]at(5){$N$}; 
		\node[above,xshift=6,yshift=5]at(8){$Leg$}; 
		\node[above,xshift=5,yshift=5]at(12){$L$};
		\draw[solid](5)circle(.05cm);
		\draw[solid](0)circle(.15cm);
		\node[right]at (11) {$W_E\left(\btn\right),$};
		\node[right,yshift=-10]at (11) {$\pi\left(\tau^n\right)-e,$};
		\node[right,yshift=-20]at (11) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};
\end{tikzpicture}

\end{document}