### 2017-01-16 to-do.tex
finish introducing and justifying -e

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage[normalem]{ulem}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true,hyperfootnotes=false]{hyperref}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
PE Shocks To-Do list
\end{center}

\section{``Dynamic preferences''}
	\begin{itemize}
		\item Becker, G. S. and K. M. Murphy (1993), Simple Theory of Advertising as a Good or Bad, Quarterly Journal of Economics, 942-64.
			\begin{itemize}
				\item Their firms are my firm. Their consumers are my government
				\item Kyle's lit review talkes about `persuasive,' `informative,' and `complementary' views of advertisting. This is complementary.
				\item $e$ and $\tau$ enter objective function. Gov't is sophisticated, can make commitment that alters $e$ (i.e. behavior of another actor). But $CS + \ga(e)PS + \ldots$, just like $u(X,Y,A)$; says consumer could work to pass law about number of ads on TV each hour.
				\item Use newly drawn extensive form to explain this, lays out when welfare is realized
				\item Addresses James' concern about needing to be more precise about when $\ga$ is evaluated
					\begin{itemize}
						\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
					\end{itemize}
			\end{itemize}
		\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
				\begin{itemize}
					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
					\item They don't like gov't choosing its own preferences. May just need to change wording
				\end{itemize}
		\item Bob: what is fundamental structure of preferences?
		\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
			\begin{itemize}
				\item ``How the assumption of a non-unitary legislature translates into this incentive constraint. In the non-unitary legislature, if the current period incentive constraint includes any future value of lobbying effort, this would mean the current-period median legislator evaluating her current-period incentive constraint using a mixture of her own political-economy weight with those of the legislators who are median in the future in the trade agreement and trade-war scenarios. The first major step I have taken to clarify the incentive constraint is the analysis in Appendix B.1 of the unitary model. I discuss the differences between the unitary and non-unitary models in the beginning of that Appendix.''
				\item ``Note that the median legislator, whose identity is determined by the lobby's effort level $e_b$, evaluates future payoffs according to her own political economy weight, $\ga(e_b)$. Of course, depending on legislator $e_b$'s choice, either legislator $e_a$ or legislator $e_{tw}$ will be the decision maker in those future periods. But legislator $e_b$, who is the decision maker in the current period, maximizes her own welfare given the predicted behavior of future decision makers.''
			\end{itemize}
		\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
				\begin{itemize}
					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's ``expost $\ga(e)$ as effort to elect/choose ex-post decision-maker''
				\end{itemize}
								\item Unitary vs. non-unitary
					\begin{itemize}
						\item Assume the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated. At the ex-ante stage, $e=0$. If this government is also unitary, then this interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
						\item Note that this remains a non-unitary model. Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g. during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$). A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results. Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}. In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort. This distinction is mainly important in the context of the repeated game. Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker. Thus the decision maker at the time the lobby is pushing for the agreement to be broken ($e=e_b>e_a$ in order to have a median legislator who will break the agreement) is different from the decision maker during the trade war phase $(e = e_{tw})$ and the decision maker during a trade agreement phase ($e=e_a(\tau^a)$). The weight on the lobbying industry's profits changes with lobbying effort because lobbying effort determines a different median voter in the legislature or government more generally. However, when the median voter during a `break' phase is evaluating her incentive constraint, she values future tariff choices --- which will be determined by whoever is median in the future --- with her own preferences. Although she will not be the decision maker in the future, there is no reason for her to evaluate future welfare according to some other legislator's preference.
						\item If the model is interpreted as having a unitary government, there is only one decision maker. There is a fixed mapping from lobbying effort to the weight the decision maker places on the lobby's profits, and the realization of this weight changes with the realization of lobbying effort. Thus, predicting future lobbying effort, the decision maker knows what the realization of her weight on the lobby's profits will be and evaluates the incentive constraint accordingly. Expression (10) would be modified while the lobbying constraint is unchanged.
					\end{itemize}
		\end{itemize}

\section{BS2005 vs. MRC2007}
	\begin{itemize}
		\item I like the way you use Bagwell and Staiger (2005) as your benchmark.  But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.  I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.  In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.  But your mechanisms seem quite different.  They have capital mobility whereas your factors are specific to each sector.  If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.  What drives this feature in your model?  Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.  What is the relationship between lobby effort level and tariff policy?  Is effort fixed first, after which the government sets its tariff?  If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.  It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC. 
		\item DGH / MRC analogy from JIE$\_$revision
			\begin{itemize}
				\item After Prop1 is a good place to start to put references to MRC 2007

				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007. Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored. In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps. The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff. There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff. Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement. A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement. From the ex-ante point of view, strong bindings are therefore preferable. Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
			\end{itemize}
		\item MRC07
			\begin{itemize}
				\item Need case of fixed capital ($z=0$)
				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
					\begin{itemize}
						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period). Here, could be free trade but could be interior to get some rents.
					\end{itemize}		
			\end{itemize}
	\end{itemize}

\section{Lit review}
\begin{itemize}
	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
		\begin{itemize}
			\item Bagwell and Staiger 1999 AER
		\end{itemize}
	\item Have to make connection to endogenous papers clearer
		\begin{itemize}
			\item More on Coates $\&$ Ludema (2001)
			\item Check out Ben and James new paper
		\end{itemize}
	\item Paola: How do I want to sell the paper?
		\begin{itemize}
			\item In relation to what literature?
				\begin{itemize}
					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
					\item What literature \textit{doesn't}?
				\end{itemize}
			\end{itemize}
	\item Review Bown, Bagwell $\&$ Staiger
	\item Do I include papers like BS2001 (non-repeated) in lit review?
	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with
	 contributions `felonious'
	\item David DeRemer: Safeguards have dropped off in use around the world
		\begin{itemize}
			\item Look at U.S. vs. ROW (he sent papers)
			\item Wouters and Zissimos (2016 WP): Appellate Body failed to give clear direction in cases in early 2000s as to when Safeguards can be applied. See Sykes (2003).
		\end{itemize}
\end{itemize}

\section{More on escape clause / WEC}
		\begin{itemize}
			\item The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.  In fact I was itching to get to that part right from when you discussed it in the introduction.  
				\begin{itemize}
					\item The paper makes the extremely important point almost in passing.  I would be interested to see a more detailed examination of how this works.  It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.  I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
				\end{itemize}
			\item `Withdrawal of equivalent concessions' (WEC).
				\begin{itemize}
					\item Zissimos (2007): a government \textbf{`chooses the severity of its own punishment'} by the extent of its initial deviation. I'm wondering whether your framework offers an answer for why WEC made sense as an approach to punishment.
					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
						\begin{itemize}
							\item \textbf{Think about how WEC may mitigate LOBBY'S incentives through reaction of gov't}
							\item Ben doesn't have any need for escape; have to put WEC into my framework
							\item Government feels $\ga$ the same whether it's elevated because of $s$ or $e$
							\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
							\item WEC should work against TOT, not PE shock?
						\end{itemize}
					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated.
					\item I say I'm going to do this on Page 27. Need to decide and either do it or take this out.
				\end{itemize}
			\item Puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round, even though no retaliation for 3 yrs
				\begin{itemize}
					\item ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), %\url{C:\Users\Kristy\Dropbox\Research\PEshocks\Literature\2016}
					\item Need to understand better what changed, both with safeguard and with other policy instruments
					\item Note that safeguards agreement states that safeguard can only be at level necessary to remedy injury, and must liberalize as possible
					\item \textbf{perhaps better able to respond to $\ga(e)$ before dispute settlement?}
						\begin{itemize}
							\item Lack of effective enforcement before WTO?
						\end{itemize}
				\end{itemize}
			\item What does WTO \textit{really} want? To discourage rent-seeking lobbying but allow governments to escape when there's a real shock.
				\begin{itemize}
					\item There may be legitimate lobbying to communicate about the shock, so can't look at the presence of lobbying as a sufficient statistic
				\end{itemize}
			\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
		\end{itemize}

\section{Costly state verification}
		\begin{itemize}
				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?  I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.  How does your approach relate to costly state verification?
				\item I say that I'm going to provide a result along the lines of Beshkar 2010b. Same or different from costly state verification? Which is more helpful?
				\item Need to check this for sure in new section. I don't have cost to verification.
		\end{itemize}

\section{Dynamic use constraint}
When would lobby exert effort to top up?
\begin{itemize}
	\item Do I want correlation between endogenous/exogenous parts? 
\end{itemize}

\section{Motivation}
		\begin{itemize}
			\item What it's missing: Well articulated question / paradox (I think it's just not well-articulated)
			\item Intro doesn't reflect body of paper
			\item What kinds of empirical questions might we ask with this model?
			\item What would be different because we use this model?
			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper. However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government. Is this aspect crucial for the point you want to make? If yes, it should be integrate with the previous object from the beginning. If not, it should be postponed a bit.
			\item I also find the introduction too long. You could/should split the literature review that begins on page 4 to a separate section. And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it. Again, your main point seems to get lost into the discussion instead of standing out.
			\item You need to think a bit more about the objective(s) of the paper and how best to achieve them.
			\item Intro: GH gives micro foundations. But there are real consequences of endogenous politics.
		\end{itemize}

\section{Big picture of which changes do what}
\vskip.2in
		\begin{tabular}{|l|c|c|}
			 & rigid & escape \\
			BS2005  & on-schedule (truthtelling): trivial; & need cost  \\
			 & off (repeated): optimal static if patient enough & \\
			$\ga(e)$ & on: trivial& no need, but revisit side payments\\
			 & off: mabye not b/c of lobby & \\
			$\ga(e,s)$ & ignore for now & is $s$ verifiable? def not w/$\ga(s,e)$
		\end{tabular}
\vskip.2in
\begin{itemize}
	\item Make sure this story is clear in text: I establish baseline tariff cap case, then add escape clause. We can already see a story of $\ldots$ emerge...
	\item Need to be clear when it is flexibility, enforcement, endogenous $\ga$ that screws things up. THEN need to bmake sure it comes out in paper
	\item May want to ditch repeated game part...
\end{itemize}

\section{One good idea / do too much}
		\begin{itemize}
			\item You can have one good idea in a paper, the rest get lost
			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
			\item The paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function.
		\end{itemize}

\section{Misc.}
\begin{itemize}
	\item Specialize to BS2005 when necessary (tariff cap example)
	\item Define $\ga$ before using it
	\item If add back in assumption 1, add Ethier footnote
	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
	\item Ben: I'm wondering how you view the time-frame over which your model is set.  On one hand a specific-factors set-up is normally associated with the short-run.  But on the other hand you have an infinite time horizon which suggests very long run.
	\item Clarify information structure
	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
	\item Add ref to JIE R$\&$R to fn:krw ?
	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
\end{itemize}

\section{Existence proofs?}

\newpage
Current structure of paper (August 25, 2016)
\begin{enumerate}
	\item Introduction
	\item Model
	\item Rigid Tariffs with Endogenous Political Pressure
		\begin{itemize}
			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
			\item[3.2] Self-Enforcing Trade Agreements
				\begin{itemize}
					\item[3.2.1] Repeated Game
					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
				\end{itemize}
		\end{itemize}
	\item Endogenous Political Pressure and the Escape Clause
		\begin{itemize}
			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
			\item[4.2] Side payments: spirit not upheld, but IC
			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
			\item[4.4] EC for endogenous politics
		\end{itemize}
	\item Conclusion
\end{enumerate}

\newpage
\begin{center}
\textbf{\large{For Later}}
\end{center}
\begin{enumerate}
	\item Add export lobby. 
		\begin{itemize}
			\item Extension or appendix b/c want to make comparison to existing literature.
		\end{itemize}
\end{enumerate}
\end{document}