### 2017-07-10 contests.tex
add contests.tex with lit suggestions from Dan Kovenock

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage[normalem]{ulem}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true,hyperfootnotes=false]{hyperref}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
Contests literature
\end{center}

Suggestions from Dan Kovenock, July 2017
\begin{itemize}
	\item Roberson and Kovenock survey on Contests with Multiple Battlefields, Oxford Handbook of the Economics of Peace and Conflict, Michele Garfinkel and Stergios Skaperdas eds.
	\item Snyder, with looks at both the expected sum of battlefields won objective and the majoritarian objective.
	\item Klumpp and Polborn: cases (with 7+  targets) when the equilibrium in the majoritarian case is only in nondegenerate mixed strategies.
	\item Several by Fletcher and Slutsky 
		\begin{itemize}
			\item they may have more since these, you might look and see
		\end{itemize}
	\item Alexander Robson (student of Stergios): multibattle contests with Tullock CSF. 
\end{itemize}

\end{document}