### 2018-02-25 to-do.tex
add ideas from IER referee reports to to-do.tex

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage[normalem]{ulem}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true,hyperfootnotes=false]{hyperref}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\ga}{\gamma}
\newcommand{\Ga}{\Gamma}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
PE Shocks To-Do list
\end{center}

\section{From IER Referee Reports}
\begin{itemize}
	\item Is there a proposition to support each of my main points?
	\item Make results more formal
		\begin{itemize}
			\item explicitly provide mappings from exogenous parameter regions to predictions about endogenous variables. For example, Proposition 2 simply indicates that a certain prediction ``may'' emerge without providing a detailed formal characterization.
		\end{itemize}
	\item Result 2 is interesting but not new
	\item Clarifying points
		\begin{itemize}
			\item I would recommend saying on p. 7 what it means to “form a trade agreement.” I believe at this point in the paper the meaning is a tariff binding of some kind.
			\item I would recommend formally defining the Wx function on p. 9. This is used often in the sequel, and I was unclear as to whether this function includes the lobbying cost “-e” that is found in expression (1).
			\item In expression (3) on p. 10, it would be helpful to discuss the tariff function that is embedded in the profit function. Is this the “trade agreement”? By using this function, are we assuming that political pressure is observed here? I was a little uncertain as to exactly what was being represented here.
			\item In expression (5) on p. 11, where is the expectation over s and s*? Isn’t the trade agreement formed in anticipation of possible economic shocks? Since political pressure is privately observed, should the reader understand this as a complete-information benchmark of some kind? If so, is this the same “trade agreement” tariff that was mentioned previously?
			\item On p. 14, rigid tariffs are defined as corresponding to a situation in which governments have no provision for flexibility, and the section is entitled “Rigid tariffs with Endogenous Political Pressure.” But in the first subsection (on p. 15) we are told that attention is restricted to the case of weak bindings throughout. Weak bindings seem to provide flexibility to apply lower tariffs, so they would not seem to be rigid tariffs, which makes the title to the section seem odd. Some additional clarity about terminology would be helpful here.
			\item On p. 17, I am not sure what τN in the third paragraph is. Do you have some specific pressure in mind that is suppressed? Or is this the tariff that is associated with eL? Or perhaps it was defined somewhere and I missed it. Similar remarks apply to expression (8) on p. 20. What is the definition of eN there? Is it different from eL?
			\item On p. 21, line 5, I don’t think you meant to say that the left hand side is decreasing in $\delta$, or maybe I am misunderstanding something.
		\end{itemize}
	\item Does costly state verification create threshold for endogenous $e$ even with verifiable shock conditionality?
	\item Differentiate from BS2005: they have exogenous only, so can't get my insight on 'free pass' on page 2; any paper that gets it with exogenous shocks only are fundamentally different
		\begin{itemize}
			\item I'm showing something new: that endogenizing politics in these models \textit{matters}
		\end{itemize}
	\item Is my write-up very different from the presentation? What are my main points?
	\item Clarify that it's not the $\gamma(e)$ that replaces Mitra / MRC inefficiencies; these are very complicated ways to get the government welfare function to be not everywhere increasing in lobbying effort.
		\begin{itemize}
			\item What I've done is provide a tractable way to get it
			\item May need to add more: Mitra 1999, Mitra 2002, Krishna and Mitra 2005
			\item Yes, relative to GH, mechanism by which greater lobbying activity leads government to give more weight to lobbying industry's profits is more of a black box. But GH functional form isn't very reasonable, and this is much better than what is usually done in the literature with exogenous shocks only. How would you put shocks into GH framework?
				\begin{itemize}
					\item Need to emphasize this being able to do shocks at same time as endogenous pressure
				\end{itemize}
		\end{itemize}
	\item Clarify that by long-run inefficiencies I also mean the lobbying inefficiencies of Mitra
	\item Do I really restrict attention to infinitely patient governments in Proposition 2?
\end{itemize}
\section{Existence proofs?}

\section{Dynamic use constraint}
When would lobby exert effort to top up?
\begin{itemize}
	\item Do I want correlation between endogenous/exogenous parts? 
\end{itemize}

\section{Add export lobby}
		\begin{itemize}
			\item Extension or appendix b/c want to make comparison to existing literature.
		\end{itemize}

\section{Lit review}
\begin{itemize}
	\item Have to make connection to endogenous papers clearer
		\begin{itemize}
			\item More on Coates $\&$ Ludema (2001)
			\item Check out Ben and James new paper
		\end{itemize}
	\item Paola: How do I want to sell the paper?
		\begin{itemize}
			\item In relation to what literature?
				\begin{itemize}
					\item What literature does endogenous? Coates-Ludema 2001, MRC, GH
					\item What literature \textit{doesn't}?
				\end{itemize}
			\end{itemize}
	\item Review Bown, Bagwell $\&$ Staiger
	\item Kyle: McLaren in his chapter calls the Grossman and Helpman setup with contributions `felonious'
	\item David DeRemer: Safeguards have dropped off in use around the world
		\begin{itemize}
			\item Look at U.S. vs. ROW (he sent papers)
			\item Wouters and Zissimos (2016 WP): Appellate Body failed to give clear direction in cases in early 2000s as to when Safeguards can be applied. See Sykes (2003).
		\end{itemize}
	\item Read Rosendorff $\&$ Milner (2001) [from Cristiane Carneiro at PEIO]
	\item Find MRC cites (maybe some in Ethier) for why PE shocks can't be addressed
		\begin{itemize}
			\item Bagwell and Staiger 1999 AER
		\end{itemize}
\end{itemize}

\section{Misc}
\begin{itemize}
	\item Ben: I'm wondering how you view the time-frame over which your model is set.  On one hand a specific-factors set-up is normally associated with the short-run.  But on the other hand you have an infinite time horizon which suggests very long run.
	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
\end{itemize}

\section{More on escape clause / WEC}
		\begin{itemize}
			\item The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.  In fact I was itching to get to that part right from when you discussed it in the introduction.  
				\begin{itemize}
					\item The paper makes the extremely important point almost in passing.  I would be interested to see a more detailed examination of how this works.  It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.  I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
				\end{itemize}
			\item `Withdrawal of equivalent concessions' (WEC).
				\begin{itemize}
					\item Zissimos (2007): a government \textbf{`chooses the severity of its own punishment'} by the extent of its initial deviation. I'm wondering whether your framework offers an answer for why WEC made sense as an approach to punishment.
					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
						\begin{itemize}
							\item \textbf{Think about how WEC may mitigate LOBBY'S incentives through reaction of gov't}
							\item Ben doesn't have any need for escape; have to put WEC into my framework
							\item Government feels $\ga$ the same whether it's elevated because of $s$ or $e$
							\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
							\item WEC should work against TOT, not PE shock?
						\end{itemize}
					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated.
				\end{itemize}
			\item Puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round, even though no retaliation for 3 yrs
				\begin{itemize}
					\item ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), %\url{C:\Users\Kristy\Dropbox\Research\PEshocks\Literature\2016}
					\item Need to understand better what changed, both with safeguard and with other policy instruments
					\item Note that safeguards agreement states that safeguard can only be at level necessary to remedy injury, and must liberalize as possible
					\item \textbf{perhaps better able to respond to $\ga(e)$ before dispute settlement?}
						\begin{itemize}
							\item Lack of effective enforcement before WTO?
						\end{itemize}
				\end{itemize}
			\item What does WTO \textit{really} want? To discourage rent-seeking lobbying but allow governments to escape when there's a real shock.
				\begin{itemize}
					\item There may be legitimate lobbying to communicate about the shock, so can't look at the presence of lobbying as a sufficient statistic
				\end{itemize}
			\item Does WEC help or hurt vs. grim trigger, $T$ period Nash reversion?
		\end{itemize}

\newpage
Current structure of paper (August 25, 2016)
\begin{enumerate}
	\item Introduction
	\item Model
	\item Rigid Tariffs with Endogenous Political Pressure
		\begin{itemize}
			\item[3.1] Perfect External Enforcement. Proposition 1: weak bindings and ext. enforcement imply applied tariff = binding and may use binding to encourage or restrain lobbying
			\item[3.2] Self-Enforcing Trade Agreements
				\begin{itemize}
					\item[3.2.1] Repeated Game
					\item[3.2.2] Prop 2: No ext enforcement: self enforcing implies $\tau^a \leq$ optimal binding with external enforcement
				\end{itemize}
		\end{itemize}
	\item Endogenous Political Pressure and the Escape Clause
		\begin{itemize}
			\item[4.1] Strong bindings: no cost when $\ga(e)$ only
			\item[4.2] Side payments: spirit not upheld, but IC
			\item[4.3] $\ga(s,e)$, Prop 3: lower binding never used
			\item[4.4] EC for endogenous politics
		\end{itemize}
	\item Conclusion
\end{enumerate}

\end{document}